---
title: 'Ames Housing: Data Cleaning'
author: "Richard Ryan"
date: '2022-04-07'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, paged.print = FALSE)
```

# Introduction

The Ames Housing Dataset is one of the most popular in data-science. There have been thousands of submissions on the [Kaggle website](https://www.kaggle.com/competitions/house-prices-advanced-regression-techniques) and it will be interesting to see how our analysis compares. 

There are several ways of accessing the dataset: we have a built-in version in `R` and we also have the files made available on [Kaggle](https://www.kaggle.com/competitions/house-prices-advanced-regression-techniques/data). However, the data we shall use comes instead from [this page](https://www.openintro.org/book/statdata/index.php?data=ames).

The reason for not using the built in `AmesHousing` package is because the data is slightly different from that provided by [Kaggle](https://www.kaggle.com/competitions/house-prices-advanced-regression-techniques), making a direct comparison difficult between results problematic. We have not used the [Kaggle data](https://www.kaggle.com/competitions/house-prices-advanced-regression-techniques/data) because it would not be possible to show the results from within this analysis. 

So let's begin our analysis by loading the libraries we shall need. We use the `tidyverse` framework for exploring and manipulating our data, the `janitor` package for cleaning up the names of variables, and the `modeest` package for the `mfv()` function which we shall use frequently in what follows.  

```{r}
library(tidyverse)
library(tidymodels)
library(janitor)
library(modeest)
```

Our analysis will follow a fairly standard path:

* Read in the data
* Explore the data
* Clean the data
* Engineer new features
* Build a model and predict on new data
* Evaluate the results

# Read in our data

The first step is very simple. We use the `read_csv()` function from `readr` to import the file. Using the `clean_names()` function from `janitor` will make the variables much easier to work with. We then split the file into training and test sets, using `set.seed()` to ensure our results can be easily reproduced. 

We make a copy of the `sale_price` variable from the test set, but we delete `sale_price` from the test set itself to avoid any possible data-leakage. 


```{r}
ames_raw <- read_csv("ames.csv") %>% 
   clean_names()
```

```{r}
set.seed(2022)
```

We now
```{r}
ames_splits <- initial_split(ames_raw, prop = 0.7)

ames_train <- training(ames_splits)
ames_test <- testing(ames_splits)
```

```{r}
sale_price <- ames_test %>% 
   select(sale_price)

ames_test <- ames_test %>% 
   select(-sale_price)
```

```{r}
ames <- bind_rows(ames_train, ames_test)
```

```{r}
remove(ames_raw, ames_train, ames_test, ames_splits)
```

# Exploring our data: missing values

In this section we shall examine our data for any missing values. Fist let's write a simple function, using `map_dfr()` from `purrr` to test for features whose missing values make up more than 50% of the total:

```{r}
ames %>% 
   map_dfr(function(.x) tibble(
      total = length(.x),
      total_NAs = sum(is.na(.x)),
      percent_NAs = mean(is.na(.x))),
      .id = "variable"
   ) %>% 
   filter(percent_NAs > 0.5) %>% 
   arrange(desc(percent_NAs))
```

These feature are missing so many values that we must drop them from our model. We could easily delete these four values by eye; but doing the job with code is always better:


```{r}
excess_NAs <- ames %>% 
   summarise(across(everything(), function(.x) mean(is.na(.x)))) %>%
   select_if(function(.x) .x > 0.5) %>% 
   names()

excess_NAs
```

```{r}
ames <- ames %>% 
   select(-all_of(excess_NAs))

remove(excess_NAs)
```

For all other variables we can at least attempt to impute any missing values. Again, let's use `map_dfr()` to see the scale of the problem:

```{r}
ames %>% 
   map_dfr(function(.x) tibble(
      total_NAs = sum(is.na(.x))),
      .id = "variable") %>% 
   filter(total_NAs > 0) %>% 
   arrange(desc(total_NAs)) %>% 
   print(n = Inf)
```

We can assume that an `NA` for `fireplace_qu` simply means there is no fireplace. This problem is easily fixed using the `replace_na()` function from `tidyr`:

```{r}
ames <- ames %>% 
   mutate(fireplace_qu = replace_na(fireplace_qu, "none"))
```

The sale price is meant to have missing values, so nothing needs to be done there. Imputing values for the `lot_frontage` feature is more easily done using an algorithm at the `recipe()` stage, so this feature can also be left for the time being.

What about the garage variables? 

## Garage variables

We have to treat these variables as a group, as the way in whcih we fix one will affect how we can approach the others.

For example, We are told in the [data description](https://www.kaggle.com/competitions/house-prices-advanced-regression-techniques/data) on [Kaggle](https://www.kaggle.com/) that an `NA` for `garage_type` mean that the property lacks a garage. Therefore this feature must be addressed first:

```{r}
ames <- ames %>% 
   mutate(garage_type = replace_na(garage_type, "none"))
```

Obviously any other category falling under garage can be replaced accordingly. If there is no garage, then all other `character()` variables relating to garage can be set to `none` also. 

The use of the `across()` function makes this a very straightforward task.

```{r}
ames <- ames %>% 
   mutate(across(
      where(is.character) & contains("garage"),
      function(.x) if_else(garage_type == "none", replace_na(.x, "none"), .x)
   ))
```

What missing `character()` values remain for the garage variables?

```{r}
ames %>% 
   summarise(across(
      where(is.character) & contains("garage"),
      function(.x) sum(is.na(.x))
   )) %>% 
   select_if(function(.x) .x > 0)
```

All of the remaining values can be replaced with the mode, for which we use the `mfv()` function from `modeest`:

```{r}
ames <- ames %>% 
   mutate(across(
      where(is.character) & contains("garage"),
      function(.x) replace_na(.x, mfv(.x))
   ))
```

Now we can consider the numerical properties with a garage:

```{r}
ames %>% 
   summarise(across(
      where(is.numeric) & contains("garage"),
      function(.x) sum(is.na(.x))
   ))
```

For any house without a garage we can substitute zero:

```{r}
ames <- ames %>% 
    mutate(across(
       where(is.numeric) & contains("garage"),
       function(.x) if_else(garage_type == "none", 0, .x)
   ))
```

Do any `NA`s remain for the numerical garage features?

```{r}
ames %>% 
   summarise(across(
      where(is.numeric) & contains("garage"),
      function(.x) sum(is.na(.x))
   ))
```

The `garage_yr_blt` feature can be fixed by imputing the year in which the house was built.

```{r}
ames <- ames %>% 
   mutate(garage_yr_blt = if_else(is.na(garage_yr_blt), year_built, garage_yr_blt))
```

The garage_cars and garage_area features can be replaced with the median of their respective garage_types.

```{r}
ames <- ames %>% 
   group_by(garage_type) %>% 
   mutate(across(
      where(is.numeric) & contains("garage"),
      function(.x) replace_na(.x, median(.x, na.rm = TRUE))
   )) %>% 
   ungroup()
```

## Basement Variables

Next we can consider the basement variables. Once again, this group of features contains both numeric and character variable types. We will start with the characters:

```{r}
ames %>% 
   summarise(across(
      contains("bsmt") & where(is.character),
      function(.x) sum(is.na(.x))
   ))
```

We are told that an `NA` for `bsmt_fin_type_1` means there is no basement. Therefore:

```{r}
ames <- ames %>% 
   mutate(bsmt_fin_type_1 = replace_na(bsmt_fin_type_1, "none"))
```

Once again, we can replace all the corresponding basement variables in one go:

```{r}
ames <- ames %>% 
   mutate(across(
      contains("bsmt") & where(is.character),
      function(.x) if_else(bsmt_fin_type_1 == "none", "none", .x)
   ))
```

The missing values that remain for the `character()` basement features can be replaced by the mode:

```{r}
ames <- ames %>% 
   mutate(across(
      contains("bsmt") & where(is.character),
      function(.x) replace_na(.x, mfv(.x))
   ))
```

Next we can consider the missing values of `bsmt` variables that are numeric:

```{r}
ames %>% 
   summarise(across(
      where(is.numeric) & contains("bsmt"),
      function(.x) sum(is.na(.x))
   ))
```

The first thing to do is check whether these `NA`s correspond to houses with no basement:

```{r}
ames %>% 
   filter(bsmt_fin_type_1 == "none") %>% 
   select(contains("bsmt") & where(is.numeric)) %>% 
   filter(is.na(bsmt_fin_sf_1) | is.na(bsmt_fin_sf_2) | is.na(bsmt_unf_sf) | is.na(total_bsmt_sf) | 
             is.na(bsmt_full_bath) | is.na(bsmt_half_bath)) 
```

All of the missing values belong to houses without a basement. Therefore this is a one-step fix:

```{r}
ames <- ames %>% 
   mutate(across(
      contains("bsmt") & where(is.numeric),
      function(.x) replace_na(.x, 0)
   ))
```

## Further Missing Variables

We shall next consider the masonry variables, `mas_vnr_type` and `mas_vnr_area`. As we can see, the `NA` values align here.

```{r}
ames %>% 
   select(contains("mas_")) %>% 
   filter(is.na(mas_vnr_area) | is.na(mas_vnr_type)) %>% 
   print(n = Inf)
```

We can, therefore, assume that the `mas_vnr_type` is `none` and that the `mas_vnr_area` is `0`.

```{r}
ames <- ames %>% 
   mutate(mas_vnr_type = replace_na(mas_vnr_type, mfv(mas_vnr_type)))

ames <- ames %>% 
   mutate(mas_vnr_area = replace_na(mas_vnr_area, 0))
```

The last `NA` value to replace, at least for now, is for the `electrical` feature. Again we can simply impute the mode here:

```{r}
ames <- ames %>% 
   mutate(electrical = replace_na(electrical, mfv(electrical)))
```

As we can see, the only missing values that remain belong to the `sale_price` and the `lot_frontage` features. 

```{r}
ames %>% 
   map_dfr(function(.x) tibble(
      total_NAs = sum(is.na(.x))),
      .id = "variable") %>% 
   filter(total_NAs > 0) %>% 
   arrange(desc(total_NAs)) %>% 
   print(n = Inf)
```

This is exactly as it should be. Therefore we can now consider other matters.



































