---
title: 'Ames Housing: Modelling'
author: "Richard Ryan"
date: '2022-04-19'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, paged.print = FALSE)
```

```{r}
library(tidyverse)
library(tidymodels)
library(embed)
library(bestNormalize)
library(glmnet)
library(vip)
```

We are now ready to build our model.

First we need to split into training and test data.

```{r}
ames_train <- ames %>% 
   filter(!is.na(sale_price))

ames_test <- ames %>% 
   filter(is.na(sale_price))
```

We can now set a seed and make our cross-validation folds:

```{r}
set.seed(123)
```

Our model will need tuning, so let's create data for cross-validation.

```{r}
ames_k_folds <- vfold_cv(ames_train, v = 10, repeats = 10)
```

In the `tidymodels` ecosystem, the preprocessing is done via a `recipe`, which is a series of steps that transforms our data before passing it to the model engine.

```{r}
ames_recipe <- 
   ames_train %>% 
   recipe(sale_price ~ .) %>% 
   step_rm(order, pid) %>% 
   step_log(sale_price, skip = TRUE) %>% 
   step_lencode_mixed(ms_sub_class, outcome = vars(sale_price)) %>% 
   step_lencode_mixed(neighborhood, outcome = vars(sale_price)) %>% 
   step_lencode_mixed(exterior_1st, outcome = vars(sale_price)) %>% 
   step_lencode_mixed(exterior_2nd, outcome = vars(sale_price)) %>% 
   step_ordinalscore(all_of(!!ordinal_factors)) %>% 
   step_other(all_of(!!nominal_factors), threshold = 0.05) %>% 
   step_dummy(all_of(!!nominal_factors), one_hot = TRUE) %>% 
   step_impute_knn(lot_frontage) %>% 
   step_YeoJohnson(all_numeric_predictors()) %>% 
   step_nzv(all_predictors()) %>% 
   step_corr(all_predictors()) %>% 
   step_normalize(all_numeric_predictors())
```

We shall use linear regression to predict the outcomes. More specifically we shall use an elastic net model where the values for `penalty` and `mixture` are found via extensive tuning. 

```{r}
elastic_spec <- linear_reg(
   penalty = tune(),
   mixture = tune()
) %>% 
   set_mode("regression") %>% 
   set_engine("glmnet")
```

We wrap our `recipe` and `model_spec` into a `workflow`. This isn't necessary but it make the process much simpler. 

```{r}
ames_workflow <- workflow() %>% 
   add_model(elastic_spec) %>% 
   add_recipe(ames_recipe)
```

Tuning a model can be very time consuming, so we utilise as many cores as possible.

```{r}
doParallel::registerDoParallel()
```

We now set a seed so our results are easily reproduced:

```{r}
set.seed(2022)
```

Finally we tune the hyper-parameters of our model:

```{r}
elastic_net_tune <- tune_grid(
   ames_workflow,
   resamples = ames_k_folds,
   grid = 30
)
```

We can use the `select_best` function to find the best combination of hyper-parameters: 

```{r}
elastic_net_rmse <-  elastic_net_tune %>% 
   select_best("rmse", maximize = FALSE)
```

The last thing to do before fitting our model is to update our workflow with the `elastic_net_rmse` above. We do this using the `finalize_workflow()` function:

```{r}
final_workflow <- 
   finalize_workflow(ames_workflow, elastic_net_rmse)
```

Finally we can fit our model.

```{r}
ames_model <- final_workflow %>% 
   fit(data = ames_train)
```

The `vip` package allows us a very simple way of visualising the most important features.

```{r}
ames_model %>% 
   extract_fit_engine() %>% 
   vip(geom = "col")
```
The test of any model is how well it performs on unseen data. So let's `predict()` on our test data an see how the results hold up.

```{r}
ames_preds <- ames_model %>% 
   predict(new_data = ames_test)
```

The competition on [Kaggle is evaluated using the RMSE between the predicted value and the sale price]("https://www.kaggle.com/competitions/house-prices-advanced-regression-techniques/overview/evaluation")

Therefore we bind our predictions with the `sale_price` dataframe, converting the `sale_price` variable to the logarithmic scale. We then use the `rmse()` function form `yardstick` to give us our result. 

```{r}
ames_preds <- ames_preds %>% 
   bind_cols(sale_price) %>% 
   mutate(sale_price = log(sale_price))

ames_preds %>% rmse(.pred, sale_price)
```

The [Kaggle leaderboard]("https://www.kaggle.com/competitions/house-prices-advanced-regression-techniques/leaderboard") suggests that anything less than 0.13 is a fairly good performance, but no doubt our result could be improved upon. For example, IN OUR `recipe()` we could use `step_poly()` or `step_ns()` to capture any non-linearity. We could also investigate any possible interactions between terms using `step_interact()`.











































